
\input{tf_header.tex}
\usepackage{multicol}

\begin{document}



\lecture{True or False ? (part 3)}





{\bf Instructions.}

\vskip 2mm

In this problem you will be given a few statements. For each statement you  need to 
decide if it is true or not, and justify your answer. 



\vskip 15mm


{\bf How to justify your answer.}

\vskip 2mm

\textbullet \  In order to show that a statement is false, it suffices  to give a counterexample. 
For example, consider the the statement: 

\begin{center}
\emph{The last digit of every even number is either 2, 4, or 8.}
\end{center}

To show that this statement is false, it is enough to point out that, for example, 10 is an even number, 
but its last digit is 0. 

\vskip 2mm

\textbullet \  In order to show that a statement is true, you need to 
provide a reasoning explaining why it is true in all instances. Giving one example when it is true will 
not suffice, since the statement may not work in some other cases.  For example, consider the the statement: 

\begin{center}
\emph{If $n$ is an even number then $n+2$ is also an even number.}
\end{center}

You can justify that this is true as follows. Even numbers are integers which are multiples of $2$. 
If $n$ is even then $n = 2m$ for some integer $m$. Then $n+2 = 2m+2 = 2(m+1)$, 
which shows that $n+2$ is even.  


\hfill

\begin{cbox}[Note]
This problem will not be collected or graded. However, problems of this type will appear on  exams
in this course. Sample solutions are provided at the end. 

\end{cbox}


%%%%%%%%%%%
\newpage
%%%%%%%%%%%

{\bf For each of the statements given below decide if it is true or false. If you decide that it is true, justify your answer. 
If you think it is false give a counterexample. }

\vskip 10mm

\benu
\item[\bf a)]  If $A$, $B$, $C$ are non-zero $2\times 2$ matrices such that $AB = AC$ then $B = C$.  
\vskip 10mm

\item[\bf b)]  If $A$ is $2\times 2$ matrix such that the matrix transformation $T_A\colon \R^2 \to \R^2$
given by $T_A({\vv}) = A\vv$ is one-to-one, then $T_A$ must be onto. 
\vskip 10mm

\item[\bf c)]  If $A$, $B$ are invertible $2\times 2$ matrices, and $B\neq -A$ then $A + B$ is also invertible. 

\vskip 10mm

\item[\bf d)]  If $A$ is a $2\times 2$ matrix then $ (AA^T)^T  = AA^T$ (where $A^T$ denotes the transpose 
of the matrix $A$). 

\eenu


%%%%%%%%%%%
\newpage
%%%%%%%%%%%

{\bf Here are solutions to the questions from the previous page. You should try to answer all 
questions by yourself before reading these solutions.}

\vskip 5mm 

{\small
\benu

\item[\bf a)] FALSE. Take for example
$$
A = 
\begin{bmatrix}
1 & 1 \\
1 &  1 \\
\end{bmatrix},
\hskip  5mm
B = 
\begin{bmatrix}
1 &  1 \\
1 &  1 \\
\end{bmatrix},
\hskip  5mm
C = 
\begin{bmatrix}
2 & 2 \\
0 &  0 \\
\end{bmatrix}
$$
Then  $AB = AC = \begin{bmatrix} 2 & 2 \\ 2 &  2 \\ \end{bmatrix}$ but $B \neq C$. 

\vskip 5mm


\item[\bf b)] TRUE.  Recall that  $T_A$ is one-to-one if and only of the matrix $A$
has a pivot position in every column, and that  $T_A$ is onto if and only of the matrix $A$
has a pivot position in every row. 

Since by assumption $T_A$ is one-to-one and since $A$ has 2 columns it follows that $A$ has 
2 pivot positions. Thus, since $A$ has two rows, there will be a pivot position in every row of $A$. 
This  means that $T_A$ is onto. 
   


\vskip 5mm

\item[\bf c)]  FALSE. Take for example 
$$
A = 
\begin{bmatrix}
1 & 0 \\
0 &  1 \\
\end{bmatrix},
\hskip  5mm
B = 
\begin{bmatrix}
0 & 1 \\
1 &  0 \\
\end{bmatrix},
$$
Both $A$ and $B$ are invertible, but the matrix $A+B = \begin{bmatrix} 1 & 1 \\ 1 &  1 \\ \end{bmatrix}$ is not. 

\vskip 5mm

\item[\bf d)] TRUE.  Recall that for any matrices $A$, $B$ we have $(AB)^T = B^T A^T$. 
Taking $B = A^T$ we obtain:
$$(AA^T)^T =  (A^T)^T A^T$$   
 Since $(A^T)^T = A$ this gives $(AA^T)^T = AA^T$. 




\eenu
}



\end{document}