
\input{tf_header.tex}
\usepackage{multicol}

\begin{document}



\lecture{True or False ? (part 4)}





{\bf Instructions.}

\vskip 2mm

In this problem you will be given a few statements. For each statement you  need to 
decide if it is true or not, and justify your answer. 



\vskip 15mm


{\bf How to justify your answer.}

\vskip 2mm

\textbullet \  In order to show that a statement is false, it suffices  to give a counterexample. 
For example, consider the the statement: 

\begin{center}
\emph{The last digit of every even number is either 2, 4, or 8.}
\end{center}

To show that this statement is false, it is enough to point out that, for example, 10 is an even number, 
but its last digit is 0. 

\vskip 2mm

\textbullet \  In order to show that a statement is true, you need to 
provide a reasoning explaining why it is true in all instances. Giving one example when it is true will 
not suffice, since the statement may not work in some other cases.  For example, consider the the statement: 

\begin{center}
\emph{If $n$ is an even number then $n+2$ is also an even number.}
\end{center}

You can justify that this is true as follows. Even numbers are integers which are multiples of $2$. 
If $n$ is even then $n = 2m$ for some integer $m$. Then $n+2 = 2m+2 = 2(m+1)$, 
which shows that $n+2$ is even.  


\hfill

\begin{cbox}[Note]
This problem will not be collected or graded. However, problems of this type will appear on  exams
in this course. Sample solutions are provided at the end. 

\end{cbox}


%%%%%%%%%%%
\newpage
%%%%%%%%%%%

{\bf For each of the statements given below decide if it is true or false. If you decide that it is true, justify your answer. 
If you think it is false give a counterexample. }

\vskip 10mm

\benu
\item[\bf a)]  If $\displaystyle A = \bbm a & b \\ c & d \\ \ebm$ is a $2\times 2$ matrix which is invertible, then
the matrix $B = \bbm c & d \\ a & b \\ \ebm$ is also invertible. 

\vskip 10mm


\item[\bf b)]  If $V$ is a vector space and $T\colon V \to V$ is a linear transformation 
then $T(\zzero)  = \zzero$ (where $\zzero$ denotes the zero vector in $V$). 


\vskip 10mm

\item[\bf c)]  If $W$ be a vector space and  $V_{1}$, $V_{2}$ are subspaces of $W$, then $V_{1}\cup V_{2}$
is also a subspace of $W$. Here $V_{1}\cup V_{2}$ denotes the union of $V_{1}$ and $V_{2}$, i.e. the set of 
vectors which belong to  either $V_{1}$ or $V_{2}$ (or to both). 

\vskip 10mm

\item[\bf d)]  If $W$ be a vector space and  $V_{1}$, $V_{2}$ are subspaces of $W$, then $V_{1}\cap V_{2}$
is also a subspace of $W$. Here $V_{1}\cap V_{2}$ denotes the intersection of $V_{1}$ and $V_{2}$, i.e. the set of 
vectors which belong both  $V_{1}$ and $V_{2}$.  

\eenu


%%%%%%%%%%%
\newpage
%%%%%%%%%%%

{\bf Here are solutions to the questions from the previous page. You should try to answer all 
questions by yourself before reading these solutions.}

\vskip 5mm 

{\small
\benu


\item[\bf a)] TRUE.  The matrix $B$ is obtained from $A$ by interchanging rows, so $\det B = -\det A$. 
Since $A$ is invertible we have $\det A\neq 0$. Therefore $\det B \neq 0$, which shows that $B$ is 
invertible. 


\vskip 5mm

\item[\bf b)] TRUE.  We have:
$$
T(\zzero) = T(\zzero + \zzero) = T(\zzero) + T(\zzero)
$$
Subtracting $T(\zzero)$ from both sides we obtain $\zzero = T(\zzero)$.






\vskip 5mm

\item[\bf c)]  FALSE. Take for example $W = \R^{2}$. Let $V_{1}$ be the subspace consisting of all vectors
of the form $\bbm a \\ 0 \ebm$, where $a$ is an arbitrary number. Let $V_{2}$ be the subspace consisting 
of all vectors of the form $\bbm 0 \\ b \ebm$, where again $b$ is an arbitrary number. 
We have $\ee_{1} = \bbm 1 \\ 0 \ebm \in V_{1}$  and 
 $\ee_{2} \in \bbm 0 \\ 1 \ebm \in  V_{2}$. This gives $\ee_{1}, \ee_{2}\in V_{1}\cup V_{2}$. However 
 $\ee_{1} + \ee_{2} =  \bbm 1 \\ 1 \ebm$ is not in $V_{1}\cup V_{2}$, since this vector is neither in 
 $V_{1}$ nor in $V_{2}$. This shows that $V_{1}\cup V_{2}$ is not a subspace of $\R^{2}$.

\vskip 5mm

\item[\bf d)] TRUE.  To show this we need to check that 
\benu
\item[(1)] $\zzero\in V_{1}\cap V_{2}$
\item[\ (2)] if $\uu, \vv \in V_{1}\cap V_{2}$ then $\uu + \vv \in V_{1}\cap V_{2}$
\item[\ (3)] if $\uu \in V_{1}\cap V_{2}$ and $c\in \R$ then $c\uu  \in V_{1}\cap V_{2}$.
\eenu

Checking (1): Since $V_{1}$, $V_{2}$ are subspaces, we have $\zzero\in V_{1}$ and $\zzero\in V_{2}$, so 
$\zzero \in V_{1}\cap V_{2}$. 

\vskip 3mm


Checking (2):  Since $\uu, \vv \in V_{1}\cap V_{2}$, it means that $\uu, \vv \in V_{1}$  and 
$\uu, \vv \in V_{2}$. Since $V_{1}$ and $V_{2}$ are subspaces this gives that $\uu + \vv \in V_{1}$
and $\uu + \vv \in  V_{2}$. This means that $\uu + \vv \in V_{1}\cap V_{2}$. 

\vskip 3mm

Checking (3):  Since $\uu \in V_{1}\cap V_{2}$, it means that $\uu  \in V_{1}$  and 
$\uu \in V_{2}$. Since $V_{1}$ and $V_{2}$ are subspaces this gives that $c\uu \in V_{1}$
and $c\uu  \in  V_{2}$ for any real number $\R$. This means that $c\uu  \in V_{1}\cap V_{2}$. 

\eenu
}



\end{document}