
\input{tf_header.tex}
\usepackage{multicol}

\begin{document}



\lecture{True or False ? (part 5)}





{\bf Instructions.}

\vskip 2mm

In this problem you will be given a few statements. For each statement you  need to 
decide if it is true or not, and justify your answer. 



\vskip 15mm


{\bf How to justify your answer.}

\vskip 2mm

\textbullet \  In order to show that a statement is false, it suffices  to give a counterexample. 
For example, consider the the statement: 

\begin{center}
\emph{The last digit of every even number is either 2, 4, or 8.}
\end{center}

To show that this statement is false, it is enough to point out that, for example, 10 is an even number, 
but its last digit is 0. 

\vskip 2mm

\textbullet \  In order to show that a statement is true, you need to 
provide a reasoning explaining why it is true in all instances. Giving one example when it is true will 
not suffice, since the statement may not work in some other cases.  For example, consider the the statement: 

\begin{center}
\emph{If $n$ is an even number then $n+2$ is also an even number.}
\end{center}

You can justify that this is true as follows. Even numbers are integers which are multiples of $2$. 
If $n$ is even then $n = 2m$ for some integer $m$. Then $n+2 = 2m+2 = 2(m+1)$, 
which shows that $n+2$ is even.  


\hfill

\begin{cbox}[Note]
This problem will not be collected or graded. However, problems of this type will appear on  exams
in this course. Sample solutions are provided at the end. 

\end{cbox}


%%%%%%%%%%%
\newpage
%%%%%%%%%%%

{\bf For each of the statements given below decide if it is true or false. If you decide that it is true, justify your answer. 
If you think it is false give a counterexample. }

\vskip 10mm

\benu
\item[\bf a)]  If a set $\{\vv_{1}, \vv_{2}\}$ is a basis of $\R^{2}$ then the set $\{ \vv_{1}, \vv_{1} + \vv_{2}\}$
is also a basis of $\R^{2}$. 


\vskip 10mm


\item[\bf b)]  If  $S = \{ \vv_{1}, \vv_{2}, \vv_{3}\}$ is set of vectors in a vector space $V$ such that 
$\vv_{1}$  is the zero vector, then  $S$ cannot be a basis of $V$. 


\vskip 10mm

\item[\bf c)]  If $V$ and $W$ are subspaces of $\R^{2}$ and $\dim V = \dim W$ then $V = W$. 


\vskip 10mm

\item[\bf d)]  If $V$ and $W$ are subspaces of $\R^{2}$ and $\dim V = \dim W = 2$ then $V = W$. 


\eenu


%%%%%%%%%%%
\newpage
%%%%%%%%%%%

{\bf Here are solutions to the questions from the previous page. You should try to answer all 
questions by yourself before reading these solutions.}

\vskip 5mm 

{\small
\benu


\item[\bf a)] TRUE.  If a  set $\{ \vv_{1}, \vv_{2} \}$ is a basis of $\R^{2}$  then the matrix 
$\bbm \vv_{1} & \vv_{2}\\ \ebm$ is invertible, i.e. if $\det \bbm \vv_{1} & \vv_{2}\\ \ebm \neq 0$. 
Since
$$ \det \bbm \vv_{1} & \vv_{1}  + \vv_{2} \\  \ebm =  \det \bbm \vv_{1} & \vv_{2} \\ \ebm  \neq 0$$
we obtain that $\bbm\vv_{1} & \vv_{1}  + \vv_{2} \\ \ebm$ in an invertible matrix, which shows that the set 
$\{ \vv_{1}, \vv_{1} + \vv_{2}\}$ is a basis of $\R^{2}$. 


\vskip 5mm

\item[\bf b)]  TRUE.  In order to be a basis the set $\{ \vv_{1}, \vv_{2}, \vv_{3} \}$ must be 
linearly independent, i.e. the equation
$$x_{1}\vv_{1} + x_{2}\vv_{2} + x_{2}\vv_{3} = \zzero$$
must have only one, trivial solution. However, if $\vv_{1} = \zzero$ this equation has infinitely 
many solutions, since we can take $x_{1}$ to be any number, and $x_{2} = 0$, $x_{3} = 0$. 





\vskip 5mm

\item[\bf c)]  FALSE. Take e.g.
$$V = \Span\left( \bbm 1 \\ 0  \ebm\right)\hskip 5mm W= \Span\left( \bbm 0 \\ 1  \ebm\right)$$
Then $\dim V = \dim W = 1$, but $V \neq W$. 


\vskip 5mm

\item[\bf d)] TRUE. Recall that if $W$ is a subspace of a vector space $V$ and $\dim V = \dim W$ then 
$V = W$. In our case $\dim V = \dim W = \dim \R^{2}$ which gives $V = W = \R^{2}$. 
\eenu
}



\end{document}