
\input{tf_header.tex}
\usepackage{multicol}

\begin{document}



\lecture{True or False ? (part 6)}





{\bf Instructions.}

\vskip 2mm

In this problem you will be given a few statements. For each statement you  need to 
decide if it is true or not, and justify your answer. 



\vskip 15mm


{\bf How to justify your answer.}

\vskip 2mm

\textbullet \  In order to show that a statement is false, it suffices  to give a counterexample. 
For example, consider the the statement: 

\begin{center}
\emph{The last digit of every even number is either 2, 4, or 8.}
\end{center}

To show that this statement is false, it is enough to point out that, for example, 10 is an even number, 
but its last digit is 0. 

\vskip 2mm

\textbullet \  In order to show that a statement is true, you need to 
provide a reasoning explaining why it is true in all instances. Giving one example when it is true will 
not suffice, since the statement may not work in some other cases.  For example, consider the the statement: 

\begin{center}
\emph{If $n$ is an even number then $n+2$ is also an even number.}
\end{center}

You can justify that this is true as follows. Even numbers are integers which are multiples of $2$. 
If $n$ is even then $n = 2m$ for some integer $m$. Then $n+2 = 2m+2 = 2(m+1)$, 
which shows that $n+2$ is even.  


\hfill

\begin{cbox}[Note]
This problem will not be collected or graded. However, problems of this type will appear on  exams
in this course. Sample solutions are provided at the end. 

\end{cbox}


%%%%%%%%%%%
\newpage
%%%%%%%%%%%

{\bf For each of the statements given below decide if it is true or false. If you decide that it is true, justify your answer. 
If you think it is false give a counterexample. }

\vskip 10mm

\benu


\item[\bf a)]  If $A$ is a $2\times 4$ matrix then  $\rank(A) \leq 2$.  


\vskip 10mm


\item[\bf b)] If ${\mathcal B} = \{\bb_{1}, \bb_{2}, \bb_{3}\}$ is an orthogonal basis of $\R^{3}$, 
and $\vv\in \R^{3}$ is a vector orthogonal to both $\bb_{1}$ and $\bb_{2}$, then $\vv = c\bb_{3}$ for 
some scalar $c$. 


\vskip 10mm

\item[\bf c)]  If $V$ is a subspace of $\R^{3}$ and $\ww\in \R^{3}$ is a vector such that 
$\proj_{V}\ww = \frac{1}{2}\ww$ then $\ww$ must be the zero vector.  


\vskip 10mm


\item[\bf d)]  If $\uu_{1}$, $\uu_{2}$, $\uu_{3}$ are vectors in $\R^{3}$ such that $\uu_{1}$ is orthogonal 
to $\uu_{2}$ and $\uu_{2}$ is orthogonal to $\uu_{3}$, then $\uu_{1}$ must be othogonal to $\uu_{3}$ 




\eenu


%%%%%%%%%%%
\newpage
%%%%%%%%%%%

{\bf Here are solutions to the questions from the previous page. You should try to answer all 
questions by yourself before reading these solutions.}

\vskip 5mm 

{\small
\benu


\item[\bf a)]  TRUE. Recall that $\rank(A) = \dim \Col(A)$, and that $\dim \Col(A)$  is equal to the 
number of pivot columns of $A$. Since the matrix $A$ has two rows, it can have at most two 
pivot positions (one in each row), and so it has at most two pivot columns. 

\vskip 5mm


\item[\bf b)]  TRUE. Let 
$$[\vv]_{\mathcal B}  = 
\bbm
c_{1}\\
c_{2}\\
c_{3} \\ 
\ebm
$$
be the coordinate vector of $\vv$ relative to the basis $\mathcal B$. Then 
$$\vv = c_{1} \bb_{1} + c_{2}\bb_{2} + c_{3}\bb_{3}$$
Also, since the basis $\mathcal B$ is orthogonal, we have $c_{i} = \displaystyle\frac{\bb_{i}\cdot \vv}{\bb_{i}\cdot \bb_{i}}$
for $i=1, 2, 3$. By assumption $\bb_{1}\cdot \vv = \bb_{2}\cdot \vv = 0$, so $c_{1} = c_{2} = 0$. This gives 
$\vv = c_{3}\bb_{3}$. 



\vskip 5mm

\item[\bf c)] TRUE.  Recall that the vector $\ww - \proj_{V}\ww$  is orthogonal to the subspace 
$V$. Since $ \proj_{V}\ww\in V$ this means in particular that $\ww - \proj_{V}\ww$ is orthogonal 
to $ \proj_{V}\ww$. If $\proj_{V}\ww = \frac{1}{2}\ww$ then $\ww - \proj_{V}\ww =  \frac{1}{2}\ww$, 
so we obtain that $ \frac{1}{2}\ww$ is orthogonal to itself. However, the only vector orthogonal to 
itself is the zero vector. This gives  $ \frac{1}{2}\ww = \zzero$, and so $\ww = \zzero$. 


\vskip 5mm

\item[\bf d)]  FALSE.  Take e.g. 
$$
\uu_{1} = 
\bbm
1 \\
0 \\
0 \\
\ebm,
\hskip 5mm
\uu_{2} = 
\bbm
0 \\
1 \\
0 \\
\ebm,
\hskip 5mm
\uu_{3} = 
\bbm
1 \\
0 \\
1 \\
\ebm
$$


\eenu
}



\end{document}